---
title: "Continous Assessment 1"
subtitle: ' Ciaran Darcy,  D10123132, d10123132@mydit.ie'
output:
  html_notebook: default
  pdf_document: default
  word_document: default
---

Part One
```{r}
prob <- 0.15
Sample <- 7
```

**Q1** *Find the expected number of people that will develop a Condition*

Knowing that
$$E(x)=np$$

```{r}
Binomial_expected <- function(prob, size) {
  E=size*prob
  return(E)
}

Expected_value=Binomial_expected(prob, Sample)
Expected_value
```

**Q2** *Find the probability that exactly 4 smokers of the sample will develop a severe lung condition.*

The easiest way to calculate this manually is with the **Probability Mass Function** forumla:
$$\Pr(X = x) = f(x) = \binom{n}{x} p^x (1-p)^{n-x},$$
which is the same as the *pbinomial(x = x, n=sample, p=prob)*
```{r}
dbinom(4,size=Sample, p=prob)
```
There is a way to verify that this is the correct value: We can use the **Cumulative Distribution Function** formula and the axiom
$$F(x) = \Pr(X \leq x) = \sum_{i = 0}^{\lfloor{x}\rfloor}\binom{n}{i} p^i (1-p)^{n-i}$$
$$f(x)=F(x)-F(x-1)$$
so in this case we can use *pbinom(x = x, n=sample, p=prob)*`
```{r}
pbinom(4,size=Sample, prob = prob)-pbinom(3, size=Sample, prob=prob)
```
Since both give the same answer we can be confident that this is the correct number

**Q3** *Find the probability that more than 2 smokers of the sample will develop a severe lung condition.*

Again we have 2 ways to find this:
One uses the fact that 
$$\Pr(X > x) = 1-F(x)$$
So we can find the CDF and take one from it 

```{r}
greater3<-1-pbinom(2, size=Sample, prob=prob)
greater3

```
The second, more computationally intensive method involves 
summing the PMFs together 

$$Pr(X > x) =  \sum_{i = x+1}^{Max Size} f(i) $$
```{r}
sum(dbinom(3:Sample, size = Sample, prob=prob))

```
Again they match 

**Q4** *Find the probability that at least 3 but less than 6 smokers of the sample will develop a severe lung condition.*

In this case the we want 
$$Pr(3 \le x < 6 )$$
First thing to note is that 
$$Pr(x<X)= F(x-1)$$
Secondly we know that 
$$Pr(X\le x \le Y ) = F(Y)-F(X-1)$$
so we can use
```{r}
pbinom(5, size=Sample, prob=prob) - pbinom(2, size=Sample, prob = prob)

```
or again we can sum the probabilities being careful to note that it goes to 5 because the question explicity says less than
```{r}
between_3and6 <- sum(dbinom(3:5, size = Sample, prob=prob))
between_3and6

```
**Q5**  *Find the conditional probability that no more than 5 smokers of the sample will develop a severe lung condition given that more than 3 smokers of the sample will develop a severe lung condition.*

For this we need to make use of a number of function and capabilities and axioms

Firstly the conditional probability forumla

$$\Pr(A | B) = \frac{\Pr(A \cap B)}{\Pr(B)}.$$
so we can infer that 

$$\Pr(X\le 5 | X > 3) = \frac{\Pr(3 \le X < 6 ))}{\Pr(X > 3)}.$$

As we have already calcluated both of these all we need to do is apply the formula
```{r}
between_3and6/greater3

```
**Q6** *Create a barplot of the PMF of the random variable representing the number of smokers of the sample who will develop a severe lung condition.*


